---
title: "Data Analysis"
author: "Katy Miles"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lme4)
library(lmerTest)
library(emmeans)
library(grid)
library(gridExtra)
library(ggfortify)
library(broom.mixed)
library(knitr)
library(ggpubr)
```

## Importing Data

```{r}
# Read in data without extra label row
data = read.csv("../Input/longData.csv")
```

## Cleaning Data

```{r}
# Create a new treatment column for analysis
data = data %>% mutate(Treatment = paste0(`Culture.Variable`, " ", 
                           `Pre.acidification`),
                       Milk.Batch = dense_rank(Milk.Protein))

```

## Summary Statistics

```{r}
# Get mean, sd, min, and max for all response variables
summary_df = data %>%
  group_by(Treatment, Response) %>%
  summarise(mean = mean(value), sd = sd(value),
            min = min(value), max = max(value), 
            Grouping = Grouping) %>%
  unique()

write.csv(summary_df, "../Output/summary_stats.csv", row.names = FALSE)
```

## Plotting

```{r}
# Create a function with the response as an input variables for plotting
summary_plot = function(y) {
  select_data = data %>%
    filter(Response == y)
  
  ggplot(select_data) + 
  geom_jitter(aes(x = Treatment, 
                 y = value,
                 color = Treatment),
                 alpha = 0.7, width = 0.15) +
  ylab(y) +
  xlab("") + 
  theme_classic() 
}

# Create plots for all response variables
response = unique(data$Response)
plots = map(response, summary_plot)
names(plots) = data$Grouping[1:42]

# Save all plots in pdf (one plot per page)
all_plots = ggexport(plots, filename = "../Output/allPlots.pdf")

# Process plots
plot_page1 = ggarrange(plotlist = plots[1:6],
                        nrow = 2, ncol = 3, common.legend = TRUE)

plot_page1 = annotate_figure(plot_page1,
                             top = text_grob("Process Data", 
                                       face = "bold", size = 14))

ggexport(plot_page1, filename = "../Output/processData_plots.pdf")


# Biochemical plots
plot_page2 = ggarrange(plotlist = plots[7:23],
                        nrow = 3, ncol = 3, common.legend = TRUE)

ggexport(plot_page2, filename = "../Output/biochemicalData_plots.pdf")

# Sensory plots
plot_page3 = ggarrange(plotlist = plots[24:42],
                        nrow = 3, ncol = 3, common.legend = TRUE)

ggexport(plot_page3, filename = "../Output/sensoryData_plots.pdf")


# PCA plots
plot_pca = function(group, df) {
  df_subset = df %>% 
    filter(Grouping == group)
  wide_data = pivot_wider(df_subset,  names_from = Response)
  dependent_var = as.matrix(wide_data[,12:ncol(wide_data)])
  pca_res = prcomp(dependent_var,
                   scale. = TRUE)
  autoplot(pca_res, wide_data, colour = "Treatment") + 
    theme_classic() + 
    ggtitle(group) + 
    theme(plot.title = element_text(hjust = 0.5))
}

groups = c("Process Data", "Biochemical Data", "Sensory Data")
pca_plots = map(groups, plot_pca, df = data)

# Save pca plots as pdf for each group
ggexport(pca_plots, filename = "../Output/pcaPlots.pdf")
```

## Analysis

```{r}
# Create a function for fitting a mixed effect model
# with response variables as input for automation
fit_model = function(response) {
  select_data = data %>%
    filter(Response == response)
  model = lmer(value ~ Treatment + (1|Milk.Batch),
               data = select_data)
  return(model)
  
}

# Fit models for all response variables
model_list = map(response, fit_model)
names(model_list) = response
summary_list = map(model_list, tidy)

# Get a list of p-values
plist = as.data.frame(t(unlist(summary_list))) %>%
  select(matches("p.value")) %>%
  t() %>%
  na.omit()

plist = as.data.frame(cbind(c(rep("Process Data", 18),
              rep("Biochemical Data", 51),
              rep("Sensory Data", 57)), plist))

colnames(plist) = c("Group", "pval")

# Adjust p-values with FDR
plist_adjusted = c()
for (group in c("Process Data", 
                  "Biochemical Data",
                  "Sensory Data")) {
  p_group = as.matrix(filter(plist, Group == group))
  p = p.adjust(p_group[,2], method = "fdr")
  plist_adjusted = append(plist_adjusted, p)
}

# Create a df to compare adjusted and unadjusted pvals
pval_df = merge(as.data.frame(plist), as.data.frame(plist_adjusted),
           by="row.names") %>%
  mutate("Sig" = ifelse(plist_adjusted <= 0.05, "Yes", "No"))

colnames(pval_df) = c("response", "Grouping",
"pval_raw", "pval_adj", "sig")

write.csv(pval_df, "../Output/pval_df.csv", row.names = FALSE)

# Get pairwise comparisons for all models
emmeans_list = map(model_list, emmeans, ~Treatment)
pair_list = map(emmeans_list, contrast, "pairwise")

plot_emmeans = function(input) {
  pwpp(input) + 
    xlab("") + 
    ylab("")
}
# Plot emmeans output
emmeans_plots = map(emmeans_list, plot_emmeans)


#Create pdf of plot grids
pdf("../Output/processData_emmeans.pdf")
# Create grids for each group (3 total groups)
cowplot::plot_grid(plotlist = emmeans_plots[1:6])
dev.off()

#Create pdf of plot grids
pdf("../Output/biochemicalData_emmeans.pdf")
# Create grids for each group (3 total groups)
cowplot::plot_grid(plotlist = emmeans_plots[7:23])
dev.off()


# Create grids for each group (3 total groups)
plot = cowplot::plot_grid(plotlist = emmeans_plots[24:42])
ggsave("../Output/sensoryData_emmeans.pdf", )
```

When looking at the FDR adjusted p-values, we see that there are two significant coefficients, one for the difference between the mean coagulation time of DVI No and Bulk and one for the mean Leucine of DVI Yes and Bulk. When looking at the contrasts, we also see that there is a significant difference between the mean coagulation time of DVI No and DVI Yes. No other significant differences were found with the FDR adjustment. 

## Additional Analysis
```{r}

# MANOVA
subset_data = filter(data, Grouping == "Process Data")
wide_data = pivot_wider(subset_data,  names_from = Response)
dependent_var = as.matrix(wide_data[,12:ncol(wide_data)])
anova_list = manova(dependent_var ~
                      factor(wide_data$Treatment) +
                      factor(wide_data$Milk.Protein))
sum = summary.aov(anova_list)

p = c()
for (i in seq(1, length(sum))) {
  p = append(p, sum[[i]][1:2,5])
}

plist_adjusted = p.adjust(p, "bonferroni")
pval_df = merge(as.data.frame(p), 
                as.data.frame(plist_adjusted), 
                by="row.names") %>%
  mutate("Sig" = ifelse(plist_adjusted <= 0.05, "Yes", "No"))



# Check MANOVA assumptions
```
