---
title: "Data Analysis"
author: "Katy Miles"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lme4)
library(lmerTest)
library(emmeans)
library(grid)
library(gridExtra)
library(ggfortify)
library(broom.mixed)
library(knitr)
library(ggpubr)
```

## Importing Data

```{r}
# Read in data without extra label row
data = read.csv("../Input/longData.csv")
```

## Cleaning Data

```{r}
# Create a new treatment column for analysis
data = data %>% mutate(Treatment = paste0(`Culture.Variable`, " ", 
                           `Pre.acidification`),
                       Milk.Batch = dense_rank(Milk.Protein))

```

## Summary Statistics

```{r}
# Get mean, sd, min, and max for all response variables
summary_df = data %>%
  group_by(Treatment, Response) %>%
  summarise(mean = mean(value), sd = sd(value),
            min = min(value), max = max(value), 
            Grouping = Grouping) %>%
  unique()

write.csv(summary_df, "../Output/summary_stats.csv", row.names = FALSE)
```

## Plotting

```{r}
# Create a function with the response as an input variables for plotting
summary_plot = function(y) {
  select_data = data %>%
    filter(Response == y)
  
  ggplot(select_data) + 
  geom_jitter(aes(x = Treatment, 
                 y = value,
                 color = Treatment),
                 alpha = 0.7, width = 0.15) +
  ylab(y) +
  xlab("") + 
  theme_classic() 
}

# Create plots for all response variables
response = unique(data$Response)
plots = map(response, summary_plot)
names(plots) = data$Grouping[1:42]

# Save all plots in pdf (one plot per page)
all_plots = ggexport(plots, filename = "../Output/allPlots.pdf")

# Process plots
plot_page1 = ggarrange(plotlist = plots[1:6],
                        nrow = 2, ncol = 3, common.legend = TRUE)

plot_page1 = annotate_figure(plot_page1,
                             top = text_grob("Process Data", 
                                       face = "bold", size = 14))

ggexport(plot_page1, filename = "../Output/processData_plots.pdf")


# Biochemical plots
plot_page2 = ggarrange(plotlist = plots[7:23],
                        nrow = 3, ncol = 3, common.legend = TRUE)

ggexport(plot_page2, filename = "../Output/biochemicalData_plots.pdf")

# Sensory plots
plot_page3 = ggarrange(plotlist = plots[24:42],
                        nrow = 3, ncol = 3, common.legend = TRUE)

ggexport(plot_page3, filename = "../Output/sensoryData_plots.pdf")


# PCA plots
plot_pca = function(group, df) {
  df_subset = df %>% 
    filter(Grouping == group)
  wide_data = pivot_wider(df_subset,  names_from = Response)
  dependent_var = as.matrix(wide_data[,12:ncol(wide_data)])
  pca_res = prcomp(dependent_var,
                   scale. = TRUE)
  autoplot(pca_res, wide_data, colour = "Treatment") + 
    theme_classic() + 
    ggtitle(group) + 
    theme(plot.title = element_text(hjust = 0.5))
}

groups = c("Process Data", "Biochemical Data", "Sensory Data")
pca_plots = map(groups, plot_pca, df = data)

# Save pca plots as pdf for each group
ggexport(pca_plots, filename = "../Output/pcaPlots.pdf")
```

## Analysis

```{r}
# Create a function for fitting a mixed effect model
# with response variables as input for automation
fit_model = function(response) {
  select_data = data %>%
    filter(Response == response)
  model = lmer(value ~ Treatment + (1|Milk.Batch),
               data = select_data)
  return(model)
  
}

# Fit models for all response variables
model_list = map(response, fit_model)
names(model_list) = response
summary_list = map(model_list, tidy)

# Get a list of p-values (not including intercept term)
plist = as.data.frame(t(unlist(summary_list))) %>%
  select(matches("p.value")) %>%
  select(-matches("p.value1")) %>%
  t() %>%
  na.omit()

plist = as.data.frame(cbind(c(rep("Process Data", 18),
              rep("Biochemical Data", 51),
              rep("Sensory Data", 57)), plist))

colnames(plist) = c("Group", "pval")

# Adjust p-values with FDR
plist_adjusted = c()
for (group in c("Process Data", 
                  "Biochemical Data",
                  "Sensory Data")) {
  p_group = as.matrix(filter(plist, Group == group))
  p = p.adjust(p_group[,2], method = "fdr")
  plist_adjusted = append(plist_adjusted, p)
}

# Create a df to compare adjusted and unadjusted pvals
pval_df_mixed = merge(as.data.frame(plist), as.data.frame(plist_adjusted),
           by="row.names") %>%
  mutate("Sig" = ifelse(plist_adjusted <= 0.05, "Yes", "No"))

colnames(pval_df_mixed) = c("response", "grouping",
"pval_raw", "pval_adj", "sig")

pval_df_mixed = pval_df_mixed %>%
  mutate(response = stringr::str_remove(response, ".p.value2")) %>%
  mutate(response = stringr::str_remove(response, ".p.value3")) %>%
  cbind("Treatment" = rep(c("DVI No", "DVI Yes"), 42))

write.csv(pval_df_mixed, "../Output/pval_df_mixed.csv", row.names = FALSE)

# Get pairwise comparisons for all models
emmeans_list = map(model_list, emmeans, ~Treatment)
pair_list = map(emmeans_list, contrast, "pairwise")

plot_emmeans = function(input) {
  pwpp(input) + 
    xlab("") + 
    ylab("")
}
# Plot emmeans output
emmeans_plots = map(emmeans_list, plot_emmeans)


#Create pdf of plot grids
pdf("../Output/processData_emmeans.pdf")
# Create grids for each group (3 total groups)
cowplot::plot_grid(plotlist = emmeans_plots[1:6])
dev.off()

#Create pdf of plot grids
pdf("../Output/biochemicalData_emmeans.pdf")
# Create grids for each group (3 total groups)
cowplot::plot_grid(plotlist = emmeans_plots[7:23])
dev.off()


# Create grids for each group (3 total groups)
plot = cowplot::plot_grid(plotlist = emmeans_plots[24:42])
ggsave("../Output/sensoryData_emmeans.pdf", )

## Checking assumptions
resid_plots = map(model_list, base::plot)
```

When looking at the FDR adjusted p-values, we see that there are no significant coefficients.Thus, we fail to reject our Null hypothesis and conclude that we do not have enough evidence to prove that there is a difference between starter culture application technology on any of the process, sensory, or biochemical data.

## Additional Analysis
```{r}
# MANOVA
fit_manova = function(grouping, data) {
  subset_data = filter(data, Grouping == grouping)
  wide_data = pivot_wider(subset_data, names_from = Response)
  # Get a list of response variables
  dependent_var = as.matrix(wide_data[,13:ncol(wide_data)])
  # Fit model
  anova_list = manova(dependent_var ~
                        factor(wide_data$Treatment) +
                        factor(wide_data$Milk.Batch))
  # Get model summary
  sum = summary.aov(anova_list)
  return(sum)
}

manova_list = map(groups, fit_manova, data)
names(manova_list) = groups
# Create empty df to append pvals to
pval_df_manova = data.frame(matrix(ncol = 5))
colnames(pval_df_manova) = c("response", "pval_raw", "pval_adj", "sig", "grouping")

for (i in seq(1, length(manova_list))) {
  sum = manova_list[[i]]
  p = c()
  for (j in seq(1, length(sum))) {
    p = append(p, sum[[j]][1,5])
  }
  names(p) = names(sum)
  plist_adjusted = p.adjust(p, "fdr")
  
  pval_df_group = merge(stack(p), 
                  stack(plist_adjusted), 
                  by="ind") %>%
    mutate("Sig" = ifelse(values.y <= 0.05, "Yes", "No"))
  colnames(pval_df_group) = c("response", "pval_raw", "pval_adj", "sig")
  pval_df_group = cbind(pval_df_group, grouping = rep(names(manova_list)[i], nrow(pval_df_group)))
  pval_df_manova = rbind(pval_df_manova, pval_df_group)
  
}

write.csv(pval_df_manova[-1,], "../Output/pval_df_manova.csv", row.names = FALSE)

# Check MANOVA assumptions
```
